$\K$ of Kac---Moody groups, revisited

### 2017-06-17 paper.tex
Minor

\begin{lemma} \label{lm:normal-form} Any $m \times n$ integer matrix with parity $(c, p_c)$ can be transformed into a diagonal matrix $(d, p_d)$
  ($d = \mathrm{diag}(r_1, \ldots r_{\mathrm{min}(m, n)})$, $r_i | r_{i+1}$)
   by means of elementary operations of types~\eqref{item:type1}--\eqref{item:type5}.
\end{lemma}
\begin{proof} 
Recall that the classical algorithm for finding Smith normal form of a matrix involves
 operation of addition of one of the two fixed columns (say, $i$-th and $j$-th ones) to another
 so that, unless it was zero from the start, the minimum of $|c_{ki}|$ and $|c_{kj}|$ decreases strictly 
 (here $k$ is some row number and $i$, $j$ are some column numbers selected by the algorithm on one if its steps).

To adapt the classical algorithm to our situation of matrices with parities it suffices to describe
 how the modified algorithm performs this operation in the case when the selected columns have different parities.
Indeed, our set of elementary operations differs from the usual one only in that
 we do not allow adding odd multiples of odd columns to even columns.

@@

After the canonical form $(d, p_d)$ of $(\widetilde{A}, p_{\widetilde{A}})$ is found, it becomes trivial to calculate the group $\K(A, F)$.
Each zero diagonal coefficient $r_i$ appearing in $d$ produces one summand in $\K(A, F)$ which, depending on $p_d(i)$, is isomorphic to $\K(F)$ or $\Kmw(F)$, whereas
 each nonzero coefficient $r_i \neq 0$ gives one summand $\K(F)/(r_i \K(F))$ or $\Kmw(F) / ((r_i/2) I)$.

@@

In \cite{MW} M.~Westaway even conjectured that the above two properties could be used to read off from $q$ the parities of columns of $d$, and, thus, the type of summands appearing in $\K(A, F)$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correpond to our $q$ via $\nu = q^T$, $\mu = {q^{-1}}^T$).
While this characterization clearly works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for $q$ obtained via the classical algorithm for finding SNF, which ignores columns' parities.
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.
\begin{comment}
\begin{example}
To produce a counterexample, start from GCM 
 $A = \left(\begin{smallmatrix}
 2& -3& -2& 0\\
-1& 2& -2& 0\\
-2& -2& 2& -4\\
0& 0& -4& 2\end{smallmatrix}\right)$
  and write SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$
\[\widetilde{A} = \left(\begin{smallmatrix}
-1& 3& 0& 0\\
-2& 0& 2& 0\\
0& 0& 0& 0\\
0& -2& 2& 0\\
0& 0& 0& 0\\
0& 0& -4& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
1& 0& 0& 0& 0& 0\\
2& 1& 0& 0& 0& 0\\
0& 0& 0& 0& 1& 0\\
0& 1& 0& 1& 0& 0\\
0& 0& 0& 0& 0& 1\\
0& -2& 1& 0& 0& 0\end{smallmatrix}\right) \cdot \mathrm{diag}(-1,2,4,4) \cdot \left(\begin{smallmatrix}
1& -3& 0& 0\\
0& -3& 1& 0\\
0& -3& 0& 1\\
0& 1& 0& 0\\ \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last 2 columns of $q$ we deduce that columns with $1$st and $4$th columns of $d$ are odd while $2$nd and $3$rd are even.
Now take $s = t_{43}(1) \in \E(4, \ZZ)$, $t = t_{43}(-1) \ in \E(6, \ZZ)$.
Notice that the central factor in the decomposition $\widetilde{A} = (p t) \cdot (t^{-1} d s^{-1}) \cdot (s q)$ is still 
 a still diagonal matrix but neither of the above two criterions is applicable to the matrix $sq$.
\end{example}
\end{comment}
\printbibliography

>>>>

\begin{lemma} \label{lm:normal-form} Any $m \times n$ integer matrix with parity $(c, p_c)$ can be transformed into a diagonal matrix $(d, p_d)$
  ($d = \mathrm{diag}(r_1, \ldots r_{\mathrm{min}(m, n)})$, $r_i | r_{i+1}$)
   by means of elementary operations of types~\eqref{item:type1}--\eqref{item:type5}.
\end{lemma}
\begin{proof} 
Recall that the classical algorithm for finding Smith normal form of a matrix involves
 operation of addition of one of the two fixed columns (say, $i$-th and $j$-th ones) to another
 so that, unless it was zero from the start, the minimum of $|c_{ki}|$ and $|c_{kj}|$ decreases strictly. 
 Here $k$ is some row number and $i$, $j$ are some column numbers selected by the algorithm on one if its steps.

To adapt this algorithm to our situation it suffices to describe
 how the above operation is performed in the case when $i$-th and $j$-th columns have different parities.
Indeed, the only difference of our elementary operations from the usual ones is that
 we do not allow adding odd multiples of odd columns to even columns.

@@

Once the canonical form $(d, p_d)$ of $(\widetilde{A}, p_{\widetilde{A}})$ is found, it becomes trivial to calculate the group $\K(A, F)$.
Each invariant factor $r_i$ appearing in $d$ produces one summand in $\K(A, F)$ which, depending on $p_d(i)$, 
 is isomorphic to $\K(F)/(r_i \K(F))$ or $\Kmw(F) / ((r_i/2) I)$.

@@

In \cite{MW} M.~Westaway even conjectured that the above two properties could be used to read off the parities of columns of $d$ from $q$ and,
 thus, the type of summands appearing in $\K(A, F)$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correspond to our $q$ via $\nu = q^T$, $\mu = {q^{-1}}^T$).
Although, as said before, this characterization works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for $q$ obtained via the classical algorithm for finding SNF, which ignores columns' parities.
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.
\begin{comment}
\begin{example}
To produce a counterexample, start from GCM 
 $A = \left(\begin{smallmatrix}
 2& -3& -2& 0\\
-1& 2& -2& 0\\
-2& -2& 2& -4\\
0& 0& -4& 2\end{smallmatrix}\right)$
  and write SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$
\[\widetilde{A} = \left(\begin{smallmatrix}
-1& 3& 0& 0\\
-2& 0& 2& 0\\
0& 0& 0& 0\\
0& -2& 2& 0\\
0& 0& 0& 0\\
0& 0& -4& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
1& 0& 0& 0& 0& 0\\
2& 1& 0& 0& 0& 0\\
0& 0& 0& 0& 1& 0\\
0& 1& 0& 1& 0& 0\\
0& 0& 0& 0& 0& 1\\
0& -2& 1& 0& 0& 0\end{smallmatrix}\right) \cdot \mathrm{diag}(-1,2,4,4) \cdot \left(\begin{smallmatrix}
1& -3& 0& 0\\
0& -3& 1& 0\\
0& -3& 0& 1\\
0& 1& 0& 0\\ \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last 2 columns of $q$ we deduce that columns with $1$st and $4$th columns of $d$ are odd while $2$nd and $3$rd are even.
Now take $s = t_{43}(1) \in \E(4, \ZZ)$, $t = t_{43}(-1) \ in \E(6, \ZZ)$.
Notice that the central factor in the decomposition $\widetilde{A} = (p t) \cdot (t^{-1} d s^{-1}) \cdot (s q)$ is still 
 a still diagonal matrix but neither of the above two criterions is applicable to the matrix $sq$.
\end{example}
\end{comment}
\printbibliography
